\textbf{How open are hybrid journals included in transformative agreements?}

### 2024-09-18 analysis/manuscript.Rmd
Reviewer 2: Acknowledge impact of COVID-related literatue on OECD peak in 2020; and imprive discussion

Despite the rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020. 
In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years, from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all) |> pull(all)` in 2018 to `r country_bar$data |> filter(cat == "BRICS", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022. 
This resulted in an increase in market share of the BRICS area from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all, yearly_volume) |> mutate(prop = all /yearly_volume) |> pull(prop) * 100`% to `r country_bar$data |> filter(cat == "BRICS", cr_year == "2022") |> distinct(all, yearly_volume) |> mutate(prop = all /yearly_volume) |> pull(prop) * 100`% between 2018 and 2022, whereas that of the OECD area decreased from `r country_bar$data |> filter(cat == "OECD", cr_year == "2018") |> distinct(all, yearly_volume) |> mutate(prop = all /yearly_volume) |> pull(prop) * 100`% to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all, yearly_volume) |> mutate(prop = all /yearly_volume) |> pull(prop) * 100`% during the same period.
Upon closer examination, this trend can be observed across all of the three largest publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, particularly with regard to articles published in Physical Sciences journals. 
While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
Furthermore, OECD publication output in Health Sciences and Life Sciences journals stagnated after peaking in 2020.

@@

This study has practical implications for research funding and libraries. 
Of primary concern should be the observed differences across countries, particularly the relationship between socio-economic development and open access adoption.
In the last years, there was a notable increase in the number of closed access articles published in hybrid journals by authors from countries such as China and India, where transformative agreements and open access funding options were not widely available. 
In contrast, transformative agreements enabled a substantial increase in open access in OECD countries excluding the US, while their share of total articles in hybrid journals decreased. 
This imbalance, where open access uptake depends largely on a few less productive high-income countries, makes it increasingly unlikely that collective action towards a large-scale transition of hybrid journals to full open access through transformation agreements will be successful. 
Rather, it highlights the importance of ongoing evaluation and adaptation of open access strategies in view of global publishing trends.

>>>>

Despite the rise of open access across OECD countries, the overall publication output decreased sharply, dropping to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022 after peaking `r country_bar$data |> filter(cat == "OECD", cr_year == "2020") |> distinct(all) |> pull(all)` articles in 2020, which saw a massive growth in literature related to COVID-19, driven largely by researchers in the United States surpassing China and other countries [@Ioannidis_2021].
In stark contrast, the number of articles published in hybrid journals by first authors affiliated with institutions from BRICS countries increased steadily over the years, from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all) |> pull(all)` in 2018 to `r country_bar$data |> filter(cat == "BRICS", cr_year == "2022") |> distinct(all) |> pull(all)` in 2022. 
This resulted in an increase in market share of the BRICS area from `r country_bar$data |> filter(cat == "BRICS", cr_year == "2018") |> distinct(all, yearly_volume) |> mutate(prop = all /yearly_volume) |> pull(prop) * 100`% to `r country_bar$data |> filter(cat == "BRICS", cr_year == "2022") |> distinct(all, yearly_volume) |> mutate(prop = all /yearly_volume) |> pull(prop) * 100`% between 2018 and 2022, whereas that of the OECD area decreased from `r country_bar$data |> filter(cat == "OECD", cr_year == "2018") |> distinct(all, yearly_volume) |> mutate(prop = all /yearly_volume) |> pull(prop) * 100`% to `r country_bar$data |> filter(cat == "OECD", cr_year == "2022") |> distinct(all, yearly_volume) |> mutate(prop = all /yearly_volume) |> pull(prop) * 100`% during the same period.
Upon closer examination, this trend can be observed across all of the three largest publishers, although the shift towards BRICS is particularly evident in Elsevier's hybrid journal portfolio, particularly with regard to articles published in Physical Sciences journals. 
While OECD publication output in Elsevier's Physical Sciences journals declined from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` articles in 2018 to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "OECD") |> pull(all)` in 2022, BRICS output increased from `r publ_subject_output_df |> filter(cr_year == 2018, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` to `r publ_subject_output_df |> filter(cr_year == 2022, top_level == "Physical Sciences", publisher == "Elsevier", cat == "BRICS") |> pull(all)` in the same five-year period.
Furthermore, OECD publication output in Health Sciences and Life Sciences journals stagnated after peaking in 2020, which saw an increase in publications due to the impact of the global COVID-19 pandemic, in particular from the United States [@Ioannidis_2021].

@@

This study has practical implications for research funding and libraries.
Of primary concern should be the observed differences across countries, particularly the relationship between socio-economic development and open access adoption.
Between 2018 and 2022, there was a notable increase in the number of closed access articles published in hybrid journals by authors from countries such as China and India, where transformative agreements and open access funding options were not widely available.
In contrast, the introduction of country-wide transformative agreements in numerous OECD countries led to a substantial increase in open access, while their share of total articles in hybrid journals decreased.
This imbalance, whereby open access uptake depends largely on a few countries, makes it less likely that collective action towards a large-scale transition of hybrid journals to full open access through transformation agreements will succeed within the next years.
Rather, and in accordance with recent policy recommendations from the UK and Nordic countries [@Jisc_2024; @Widding_2024], it emphasises the importance of ongoing evaluation and adaptation of open access strategies in view of global publishing trends.
